A subjective Bayesian framework for synthesizing deep uncertainties in climate risk management

### 2022-03-30 papers/paper/preprint.tex
Update based on Klaus comments

Flood risk management relies on probabilistic estimates of flood hazard:

@@

    \item The National Flood Insurance Program was designed as a partnership between the federal government and local communities; communities can voluntarily join the program by adopting a floodplain ordinance based on the most up-to-date flood hazard maps provided by \gls{fema}. At a minimum, communities that choose to join must require that new development and substantially improved or damaged properties in high hazard areas be built at or above the level of a flood with a return period of one in 100 years (i.e., the 100-year flood). \citep{kousky_voucher:2014}
    \item Beyond flood insurance and disaster recovery there has been a federal retreat from flood risk management leaving local governments and individuals with most of the responsibility for flood risk management: ``the federal government recognizes its risks as a major owner and operator of infrastructure, facilities, \ldots, and insurer of properties and crops \ldots leaving states, localities, and individual agencies to make their own plans'' \citep{shi_transformative:2021}

@@

Future flood hazard is deeply uncertain

@@

    \item Nonstationarity also affects inland and pluvial flooding \citep{Milly:2008dg} and derives from a combination of climate change, land use change, and local river modifications \citep{Merz:2014gf}
    \item Federal guidelines for flood risk management that directly inform policy-making recognize the presence of nonstationarity but have failed to identify objectively valid approaches for modeling nonstationarity, \citep{Montanari:2014hl,Serinaldi:2015bq}

@@

    \item This is because where historic flood risk can be constrained by estimates, future flood risk depends on intrinsically unpredictable human decisions like greenhouse gas emissions, and is therefore deeply uncertain \citep{keller_management:2021}

@@

    \item This is related to equifinality where different models fit the historical record equally well but diverge as one looks into the future \citep{beven_equifinality:2006,DossGollin:2019}
    \item Applying classical optimization methods to problems with deep uncertainty can yield highly fragile solutions in the sense that guidance might change substantially if a different plausible model were used

@@

    \item RDM-like methods have been applied to coastal infrastructure planning by sampling \glspl{sow} from a \gls{pdf} of \gls{slr}, then identifying decisions that meet some predetermined satisficing criteria \citep{mcphail_robustness:2019} over a large fraction of possible futures \citep{sriver_sealevel:2018,garner_slrise:2018,lempert_slr:2012}

@@

    \item There are always subjective  modeling decisions, but sometimes they are opaque which makes it hard to communicate them to the end user
    \item The range you choose is also subjective! \citep{schneider_dangerous:2001,schneider_scenarios:2002}

@@

    \item Technical detail: joint probability distributions over high dimensional spaces are confusing. Bayesian analysis teaches us that posterior distributions of parameters are often highly correlated. If we neglect that by sampling over the marginal distribution of each parameter separately, we may be sampling almost entirely from regions of the parameter space that are not consistent with our data \citep[see][for a detailed illustration]{carpenter_dimensionality:2017}!

@@

    \item Many modern discussions of Bayesian philosophy \citep{jaynes_probability:2003,McElreath:2016vu,Gelman:2014tc,bernardo_bayesian:1994} also emphasize a philosophical view of probability as a language with which to reason about the unknown rather than a statement of objective truth \citep[see][for a thorough discussion of Bayesian philosophy]{gelman_philosophy:2013}. This is in many ways closer to the exploratory modeling proposed by \citet{bankes:1993} than the top-down, deterministic, fragile central planning approaches that it (and \cite{rittel:1973} and others) critiqued
    \item Bayesian decision theory was conceived as a calculus for reasoning rather than for identifying objective truth; \citeauthor{definetti_probability:1972} often said that ``probability does not exist'' \citeyear{definetti_probability:1972}. \citet{savage:1954} and \citet{ramsey_probability:2016}, among others, also viewed probability as ``subjective,'' representing the state of belief of the decision-maker.
    \item We know that choosing a single ``best'' model under-estimates uncertainty so inferences need to be combined across models \citep[\eg, as in stacking;][]{Yao:2018bu}

@@

    \item The famous phrase ``all models are wrong, but some are useful'' \citep[generally attributed to][]{box:1976} also suggests that probability distributions and predictions ought to be viewed subjectively.

@@

    \item Since, by definition, models are not ``true,'' this iterative workflow \citep{gelman_workflow:2020} aims to identify models that are useful and promote a dialog amongst stakeholders \citep{gelman_philosophy:2013}

@@

    \item House elevation in the coastal zone as a specific example of a problem where both over- or under-investment could be a problem

@@

Following \cref{fig:flowchart}, consider using $J$ \glspl{sow} $\vb{s} = \qty{s_1, s_2, \ldots, s_J}$ (box b in \cref{fig:flowchart}) to evaluate $I$ discrete decisions $\vb{x} = \qty{x_1, x_2, \ldots, x_I}$ (box a in \cref{fig:flowchart}).
For each scenario $s_j \in \vb{s}$ and decision $x_i \in \vb{x}$ we use a system model $f$ to calculate a set of metrics describing the performance of decision $x_i$ on \gls{sow} $s_j$, which we denote $u_{ij} = f(x_i, s_j)$ (box c in \cref{fig:flowchart}).
While we assume here that the \glspl{sow} and decisions are known and discrete, this approach could be  coupled to a policy search model that proposes candidate decisions $x_i$ from some decision space $\mathcal{X}$.\klaus{Is there better mathematical notation for panel (b)?}

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{bayes-rdm.pdf}
    \caption{
        Outline of the proposed decision analytic framework.
    }\label{fig:flowchart}
\end{figure}

@@

A first analytical step is to use the model in an ``exploratory'' mode.
Exploratory modeling avoids making explicit assumption  about the likelihood different \glspl{sow} and instead seeks to generate new knowledge through existence proofs, hypothesis generation, and vulnerability identification \citep{bankes:1993}.
Common applications include ``discovering'' particularly relevant scenarios  \citep{lamontagne_discovery:2018,groves_scenarios:2007} and mapping a system's response to decisions and \glspl{sow} \citep{Brown:2012kb,Poff:2015jn,Steinschneider:2015kk,sriver_sealevel:2018}.
Although exploratory modeling seeks to avoid explicit assumptions about the likelihood of different futures, subjective modeling decisions including the choice of system model, the set of candidate decisions, the criteria used to assess outcomes, and the choice of how to sample \glspl{sow} \citep[see][]{quinn_exploratory:2020} strongly influence results.

@@

Many decision analytic tools used in climate risk management can be interpreted through this lens.
A common application (to which we return in our case study) is when using probabilistic models with deeply uncertain boundary conditions.
For example, \ldots\james{Add a water resources and a flood example}
In this case, $M$ describes both the choice of model and the choice of boundary condition (\eg, choice of physical model and the \gls{rcp} scenario).
As we shall illustrate in \cref{sec:results-conditional}, this analysis models uncertainty \emph{within models} probabilistically, but models uncertainty \emph{between models} only as a.\ldots\klaus{I'm stuck, can you help me finish this sentence?}

Another application is to studies that generate an ensemble of \glspl{sow} by sampling parameters across fixed ranges.
For example, \citet{lempert_slr:2012} sample parameters describing the rate of \gls{slr} across a range of values to inform coastal adaptation.
Similarly, \citet{trindade_waterpathways:2020} checks the performance of candidate decisions against an ensemble of synthetic time series of streamflow, water demand, and other parameters by sampling parameters that transform the available data over a plausible range \citep[\ie, calculating robustness metrics; see][for details]{mcphail_robustness:2019,herman:2015}.
Since sampling over a range is equivalent to sampling from a Uniform distribution, this method makes an implicit assumption that $M$ is the product of $L$ independent Uniform distributions (one for each parameter) over the plausible ranges $\qty(\theta_{\ell,\min}, \theta_{\ell,\max}) \forall \ell \in \qty{1, \ldots, L}$:
\begin{equation}
    p(\Omega | M) = \prod_{\ell=1}^L \mathrm{Uniform} \qty(\theta_{\ell,\min}, \theta_{\ell,\max}).
\end{equation}
If this product of independent Uniform distributions truly represents a well-calibrated belief about the joint probability distribution over \glspl{sow} $p(\Omega)$ then it should be used, but analysts should be clear that, like it or not, they are making probabilistic assumptions, and should apply methods for model critique and validation \citep[see][for some directions]{gelman_workflow:2020}.

Our primary concern is not that subjective assumptions about the likelihood of different futures are wrong -- this is, almost surely, inevitable -- but that when these assumptions are opaque and presented without critique or validation they may lead to sub-optimal decisions.

@@

If the ensemble of \glspl{sow} available does not represent our subjective belief about the plausibility of future conditions, then some corrective measure should be taken.
One possibility is to discard the $\vb{s}$ and resample from some ``true'' distribution $p_\mathrm{belief}$.
In practice this is impractical because simulating $\vb{s}$ may rely on complicated models that are computationally expensive to re-run.
Consequently, climate risk management often uses the outputs of structured simulations and analyses \citep[\eg, from the Coupled Model Intercomparison Project's simulations;][]{meehl_cmip:2000}; it is neither practical nor feasible for local decision-makers to re-run climate models every time they want to sample \glspl{sow} from a different distribution.

Instead of attempting to re-sample $\vb{s}$, we instead consider re-weighting.
To do this we first project the \glspl{sow} $\vb{s} \in \Omega$ onto a low-dimensional representation, which we denote $\qty{\psi_1, \psi_2, \ldots, \psi_J} \in \Psi$.
This allows us to specify $p_\mathrm{belief}(\Psi)$ instead of $p_\mathrm{belief}(\Omega)$.
We then calculate a probabilistic weight $w_j \in \qty[0, 1]$ for each \gls{sow} $s_j$ such that $\sum_{j=1}^J w_j \psi_j$ approximates $p_\mathrm{belief}(\Psi)$.\klaus{Really we are setting the weights so that $\sum_{j=1}^J w_j \pi(s_j)$ approximates $\int_{s} p(s) \pi(s) \dd{s}$ -- is this too hand-wavey?}
We present here the 1D case.
Without loss of generality, assume the $\psi_j$ to be sorted from least to greatest so that $\psi_{j-1} \leq \psi_j$, ($j \neq 1$).
Defining $F_\mathrm{belief}(s)$ to be the cumulative distribution corresponding to $p_\mathrm{belief}$, we calculate weights as illustrated in \cref{fig:grid-sketch}:
\begin{equation}\label{eq:weights}
    w_j = \begin{cases}
        F_\mathrm{belief}\qty(\frac{1}{2}\qty[\psi_1 + \psi_2])                                                                     & j = 1     \\
        F_\mathrm{belief}\qty(\frac{1}{2}\qty[\psi_{j} + \psi_{j+1}]) - F_\mathrm{belief}\qty(\frac{1}{2}\qty[\psi_{j-1} + \psi_j]) & 1 < j < J \\
        1 - F_\mathrm{belief}\qty(\frac{1}{2}\qty[\psi_{J-1} + \psi_J])                                                             & j = J.
    \end{cases}
\end{equation}
For higher dimensional projections, this equation can be extended by partitioning the parameter space into a Voronoi diagram, and then integrating the \gls{pdf} $p_\mathrm{belief}$ over the boundaries of each region.
From a practical perspective, many simulation models round outputs, meaning that there is a possibility of having $\psi_j$ that are equal; in this case it may be helpful to add a small amount of noise to the $\psi_j$ before computing the weights.

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{grid-sketch}
    \caption{
        Schematic of \gls{sow} weighting scheme defined in \cref{eq:weights} for $J=4$.
        This method is illustrated for a hypothetical target distribution (orange line) and four samples $\psi_1, \psi_2, \psi_3, \psi_4$ (blue dots).
        As shown in \cref{eq:weights}, the weights $w_j$ (vertical blue lines) are calculated based on the cumulative distribution function of the target distribution at the halfway points $\frac{1}{2}\qty[\psi_i+\psi_{i+1}]$ (vertical dashed lines).
    }\label{fig:grid-sketch}
\end{figure}

The aim of this re-weighting framework is to integrate an ensemble of \glspl{sow} used for exploratory modeling into formal decision analysis.
As in \cref{sec:analysis-condition} we must condition on a model, but whereas that analysis conditions upon deep uncertainties, this approach synthesizes them.
It is crucially important to understand that stakeholders and experts will not, in general, agree on $p_\mathrm{belief}$ because its uncertainty is epistemic -- that is, there is no ``true'' value of $p_\mathrm{belief}$ that could be estimated if only we had better models and data.
However, reasoning using subjective probability allows\ldots\james{Finish sentence}

@@

We model a one-time decision of whether to elevate a house, and if so by how much (\cref{fig:xlrm}), for a \emph{hypothetical} house in Norfolk, VA.
For interpretability, we focus on deep uncertainty in \gls{msl} and treat other model parameters as more shallow uncertainties as shown in \cref{tab:uncertainties}.
To use the notation developed in the previous section, we consider the following.
\begin{enumerate}
    \item The decision vector $\vb{x}$ is a scalar describing how high to elevate a house ($\Delta h$); we consider $\Delta h = \qty{\SI{0}{ft}, \SI{0.25}{ft}, \ldots, \SI{12}{ft}}$ f
    \item The \glspl{sow} describe annual time series of \gls{msl} over the $T$ year house lifetime ($T=71$ generally; see \cref{tab:uncertainties}) so $\vb{s} \in \mathbb{R}^T$
    \item The system model $f$ quantifies up-front costs and lifetime expected damages, given a decision $x_i$ and \gls{sow} $s_j$, by integrating economic and engineering damage models over a probability distribution for storm surge.
\end{enumerate}
In the remainder of this section we describe data sources and treatment of \gls{slr} (\cref{sec:case-slr}), storm surge (\cref{sec:case-surge}), damages and metrics (\cref{sec:case-metrics}), and finally the subjective priors $p_\mathrm{belief}$ used to apply the re-weighting method described in \cref{sec:analysis-synthesize} to this case study (\cref{sec:case-priors}).

\begin{table}
    \centering
    \caption{
        Summary of parameters, their notation, and how their uncertainty is represented.
    }\label{tab:uncertainties}
    \small
    \begin{tabular}{l l p{3.5in}}
        \toprule
        Name             & Symbol            & Uncertainty                                                                          \\
        \midrule
        \Gls{msl}        & $\overline{y}(t)$ & Deeply uncertain: four physical models $\times$ four \acrshort{rcp} scenarios        \\
        Storm surge      & $y'(t)$           & Probabilistic: Bayesian inference on a stationary \acrshort{gev} model               \\
        Ann-max flood    & $y(t)$            & Deterministic: $y(t)=\overline{y}(t)+y'(t)$                                          \\
        Discount rate    & $1-\gamma$        & Fixed at 2.5\%                                                                       \\
        Depth-damage     & $D(h-y)$          & Deterministic: based on HAZUS model                                                  \\
        Elevation cost   & $C(\Delta h)$     & Deterministic: a piecewise linear model following \citet{zarekarizi_suboptimal:2020} \\
        Initial height   & $h_0$             & Deterministic: \SI{1}{ft} below the \gls{bfe} unless otherwise noted                 \\
        House floor area &                   & Deterministic: \SI{1500}{ft^2} unless otherwise noted                                \\
        Structural value &                   & Deterministic: \usd{200000} unless otherwise noted                                   \\
        House lifespan   & $T$               & Deterministic: 71 years unless otherwise noted                                       \\
        \bottomrule
    \end{tabular}
\end{table}

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{xlrm.pdf}
    \caption{
        Conceptual diagram of the decision problem.
        A \gls{sow} consists of a description of the uncertain factors (red).
        We model a problem with a single lever (yellow), which is how high to elevate a house ($\Delta h$).
        For each \acrshort{sow} and each value of $\Delta h$, the system model (blue) is used to calculate performance metrics (orange).
    }\label{fig:xlrm}
\end{figure}

@@

We analyze simulations of \gls{msl} at Sewells Point, VA from four probabilistic physical models using data published in \citet{ruckert_coastal:2019}.
The four models considered are (i) the BRICK model (version 0.2) with slow (``BRICK Slow'') and (ii) fast (``BRICK Fast'') ice sheet dynamics \citep{wong_brick0.2:2017}, (iii) the \citet{kopp_probabilistic:2014} model (``K14''), and (iv) the \citet{deconto_antarctica:2016} model (``DP16'').
The \citet{kopp_probabilistic:2014} and \citet{deconto_antarctica:2016} models have a ten year time step, which we linearly interpolate onto a one year time step for consistency.
These four models represent physical processes, particularly of ice sheet dynamics, in different ways, leading to diverging sensitivity of \gls{msl} to forcing.
For a discussion of these methods, their similarities, their differences, and their consistency with other analyses see \citet{ruckert_coastal:2019}, \citet{kopp_evolving:2017}, and \citet{bamber_slrise:2019}.

Estimates of nonstationary \gls{msl} also depend on anthropogenic forcing, which is itself deeply uncertain.
To sample this uncertainty, we use simulations from each physical model under four \gls{rcp} scenarios, yielding sixteen time-varying distributions of \gls{msl}.

@@

The choices of physical model and \gls{rcp} scenario jointly determine future \gls{msl} $p(\overline{y}|t)$.
\Cref{fig:lsl-evolution}(a) shows the time-varying \glspl{pdf} of \gls{msl} for three representative models.
The divergence between the the best-case (blue) and worst-case (red) models is small in the early 21st century and increases rapidly thereafter.
\Cref{fig:lsl-evolution}(b) shows the \glspl{pdf} of mean sea level in 2100 (dashed vertical line in panel (a)) under each of the sixteen models considered.
We return in \cref{sec:results-conditional} to the problem that these multiple models poses to decision makers.

@@

    \item Robustness metrics and robust decision analyses make implicit assumptions about the plausibility of future scenarios

@@

    \item Interacting, sequential decisions

>>>>

Flood risk management relies on probabilistic estimates of flood hazard:\james{Maybe broader, start generally and then give the example to flood risk management}

@@

    \item The National Flood Insurance Program was designed as a partnership between the federal government and local communities; communities can voluntarily join the program by adopting a floodplain ordinance based on the most up-to-date flood hazard maps provided by \gls{fema}. At a minimum, communities that choose to join must require that new development and substantially improved or damaged properties in high hazard areas be built at or above the level of a flood with a return period of one in 100 years (i.e., the 100-year flood).\james{Relate to annual exceedance probability} \citep{kousky_voucher:2014}
    \item Beyond flood insurance and disaster recovery there has been a federal retreat\james{Word choice heads up} from flood risk management leaving local governments and individuals with most of the responsibility for flood risk management: ``the federal government recognizes its risks as a major owner and operator of infrastructure, facilities, \ldots, and insurer of properties and crops \ldots leaving states, localities, and individual agencies to make their own plans'' \citep{shi_transformative:2021}

@@

Future flood hazard is deeply uncertain\james{Define deep uncertainty}

@@

    \item Nonstationarity affects inland and pluvial flooding \citep{Milly:2008dg} and derives from a combination of climate change, land use change, and local river modifications \citep{Merz:2014gf}
    \item Federal guidelines for flood risk management that directly inform policy-making recognize the presence of nonstationarity but have failed to\james{them's fighting words} identify objectively valid approaches for modeling nonstationarity, \citep{Montanari:2014hl,Serinaldi:2015bq}

@@

    \item This is because where historic flood risk can be constrained by estimates, future flood risk depends on intrinsically unpredictable human decisions like greenhouse gas emissions, and is therefore deeply uncertain \citep{keller_management:2021}\james{This is used below, avoid redundancy}

@@

    \item This is related to equifinality where different models fit the historical record equally\james{Add a qualifier} well but diverge as one looks into the future \citep{beven_equifinality:2006,DossGollin:2019}
    \item Applying classical optimization methods to problems with deep uncertainty can yield highly fragile solutions in the sense that guidance might change substantially if a different plausible model were used\james{Ref needed}

@@

    \item RDM-like\james{Explain what qualifies as RDM and how this relates to MORDM} methods have been applied to coastal infrastructure planning by sampling \glspl{sow} from a \gls{pdf} of \gls{slr}, then identifying decisions that meet some predetermined satisficing criteria \citep{mcphail_robustness:2019} over a large fraction of possible futures \citep{sriver_sealevel:2018,lempert_slr:2012,wong_nola:2017}

@@

    \item There are always subjective  modeling decisions, but sometimes these decisions are opaque which makes it hard to communicate them to the end user
    \item The range you choose is also subjective! \citep{schneider_dangerous:2001,schneider_scenarios:2002}\james{Broader Issue: You are either explicitly or implicitly a Bayesian in climate change research}

@@

    \item Technical detail: joint probability distributions over high dimensional spaces can be confusing. Bayesian analyses show that posterior distributions of parameters are often highly correlated.\james{Refs} If we neglect that by sampling over the marginal distribution of each parameter separately, we may be sampling almost entirely from regions of the parameter space that are not consistent with our data \citep[see][for a detailed illustration]{carpenter_dimensionality:2017}!

@@

    \item Many modern discussions of Bayesian philosophy \citep{jaynes_probability:2003,McElreath:2016vu,Gelman:2014tc,bernardo_bayesian:1994} also emphasize a philosophical view of probability as a language with which to reason about the unknown rather than a statement of objective truth \citep[see][for a thorough discussion of Bayesian philosophy]{gelman_philosophy:2013}. This is in many ways closer to the exploratory modeling proposed by \citet{bankes:1993} than the top-down, deterministic, fragile central planning approaches that it (and \cite{rittel:1973} and others) critique
    \item Bayesian decision theory was conceived as a calculus for reasoning rather than for identifying objective truth; \citeauthor{definetti_probability:1972} often said that ``probability does not exist'' \citep{definetti_probability:1972}. \citet{savage:1954} and \citet{ramsey_probability:2016}, among others, also viewed probability as ``subjective,'' representing the state of belief of the decision-maker.
    \item We know that choosing a single ``best'' model can drastically under-estimate uncertainty so inferences need to be combined across models \citep[\eg, as in stacking;][]{Yao:2018bu}

@@

    \item The famous phrase ``all models are wrong, but some are useful'' \citep[generally attributed to][]{box:1976} also suggests that probability distributions and predictions ought to be viewed subjectively.\james{Wording}

@@

    \item Since, by definition, models are not ``true,'' this iterative workflow \citep{gelman_workflow:2020} aims to identify models that are useful and promote a dialog amongst stakeholders and decisionmakers and analysts \citep{gelman_philosophy:2013}

@@

    \item House elevation in the coastal zone as a specific example of a problem where both over- or under-investment can lead to poor outcomes

@@

Following \cref{fig:flowchart}, consider using $J$ \glspl{sow} $\vb{s} = \qty{s_1, s_2, \ldots, s_J}$ (box b in \cref{fig:flowchart}) to evaluate $I$ discrete decisions $\vb{x} = \qty{x_1, x_2, \ldots, x_I}$ (box a in \cref{fig:flowchart}).
For each scenario $s_j \in \vb{s}$ and decision $x_i \in \vb{x}$ we use a system model $f$ to calculate a set of metrics describing the performance of decision $x_i$ on \gls{sow} $s_j$, which we denote $u_{ij} = f(x_i, s_j)$ (box c in \cref{fig:flowchart}).
While we assume here that the \glspl{sow} and decisions are known and discrete, this approach could be  coupled to a policy search model that proposes candidate decisions $x_i$ from some decision space $\mathcal{X}$.

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{bayes-rdm.pdf}
    \caption{
        Outline of the proposed decision-analytic framework.
        ADD SOME EXPLANATORY TEXT. What are the boxes and panels. % TODO
    }\label{fig:flowchart}
\end{figure}

@@

A first analytical step is to use the model in an ``exploratory'' mode.
Exploratory modeling avoids making explicit assumption  about the likelihood different \glspl{sow} and instead seeks to generate new knowledge through existence proofs, hypothesis generation, and vulnerability identification \citep{bankes:1993}.\james{May be redundant}
Common applications include discovering particularly relevant scenarios  \citep{lamontagne_discovery:2018,groves_scenarios:2007} and mapping a system's response to decisions and \glspl{sow} \citep{Poff:2015jn,Steinschneider:2015kk,sriver_sealevel:2018}.
Although exploratory modeling seeks to avoid explicit assumptions about the likelihood of different futures, subjective modeling decisions including the choice of system model, the set of candidate decisions, the criteria used to assess outcomes, and the choice of how to sample \glspl{sow} \citep[see][]{quinn_exploratory:2020} strongly influence results.\james{Link back to topic. In the topic, there is no link to observations. Make this clear and use this as a transition to the next section}

@@

Many decision analytic tools used in climate risk management can be interpreted through this lens.
A common application (to which we return in our case study) is when using probabilistic models with deeply uncertain boundary conditions.
For example, \ldots\james{Add a water resources and a flood example}
In this case, $M$ describes both the choice of model and the choice of boundary condition (\eg, choice of physical model and the \gls{rcp} scenario).
As we shall discuss in \cref{sec:results-conditional}, this analysis models uncertainty \emph{within models} probabilistically, but models uncertainty \emph{between models} only using a scenario approach (\ie, choices of model structures or factors such as forcings we show as cases) without any explicit probability statement \citep[see][]{wong_nola:2017}.\james{Wording}

Another application is to studies that generate an ensemble of \glspl{sow} by sampling parameters across fixed ranges.
For example, \citet{lempert_slr:2012} sample parameters describing the rate of \gls{slr} across a range of values to inform coastal adaptation.
Similarly, \citet{trindade_waterpathways:2020} checks the performance of candidate decisions against an ensemble of synthetic time series of streamflow, water demand, and other parameters by sampling parameters that transform the available data over a plausible range \citep[\ie, calculating robustness metrics; see][for details]{mcphail_robustness:2019,herman:2015}.
Since sampling over a range is equivalent to sampling from a Uniform distribution, this method makes an implicit assumption that $M$ is the product of $L$ independent Uniform distributions (one for each parameter) over the plausible ranges $\qty(\theta_{\ell,\min}, \theta_{\ell,\max}) \forall \ell \in \qty{1, \ldots, L}$:
\begin{equation}
    p(\Omega | M) = \prod_{\ell=1}^L \mathrm{Uniform} \qty(\theta_{\ell,\min}, \theta_{\ell,\max}).
\end{equation}
If this product of independent Uniform distributions truly represents a well-calibrated belief about the joint probability distribution over \glspl{sow} $p(\Omega)$ then it should be used, but analysts should be clear that, like it or not, they are making probabilistic assumptions, and should apply methods for model critique and validation \citep[see][]{gelman_workflow:2020}.

Our primary concern is not that subjective assumptions about the likelihood of different futures are wrong -- this is, almost surely, inevitable -- but that when these assumptions are opaque and presented without critique or validation they may lead to poor decisions.

@@

If the ensemble of \glspl{sow} available does not represent our subjective belief about the plausibility of future conditions, then some corrective measure should be taken.
One possibility is to discard the $\vb{s}$ and resample from some ``true'' distribution $p_\mathrm{belief}$.
In practice, this is impractical because simulating $\vb{s}$ may rely on complicated models that are computationally expensive to re-run.
Consequently, climate risk management often uses the outputs of structured simulations and analyses.
For example, many studies use outputs from the Coupled Model Intercomparison Project's simulations \citep{meehl_cmip:2000} to assess risks and design adaptation strategies.\james{Add some refs}
It is not practical or feasible for end users to re-run such complex Earth system models every time they want to sample \glspl{sow} from a different distribution.

Instead of attempting to re-sample $\vb{s}$, we instead consider re-weighting.
To do this we first project the \glspl{sow} $\vb{s} \in \Omega$ onto a low-dimensional representation, which we denote $\qty{\psi_1, \psi_2, \ldots, \psi_J} \in \Psi$.
This allows us to specify a prior belief over this reduced space $p_\mathrm{belief}(\Psi)$ instead of over the full space $p_\mathrm{belief}(\Omega)$.
We then calculate a probabilistic weight $w_j \in \qty[0, 1]$ for each \gls{sow} $s_j$ so that the weighted distribution of $\psi_j$ closely approximates $p_\mathrm{belief}(\Psi)$.

We present here the case where the $\psi_j$ are one-dimensional.
We first sort the $\psi_j$  from least to greatest so that $\psi_{j-1} \leq \psi_j$, ($j \neq 1$).
Defining $F_\mathrm{belief}(s)$ to be the cumulative distribution corresponding to $p_\mathrm{belief}$, we calculate weights as
\begin{equation}\label{eq:weights}
    w_j = \begin{cases}
        F_\mathrm{belief}\qty(\frac{1}{2}\qty[\psi_1 + \psi_2])                                                                     & j = 1     \\
        F_\mathrm{belief}\qty(\frac{1}{2}\qty[\psi_{j} + \psi_{j+1}]) - F_\mathrm{belief}\qty(\frac{1}{2}\qty[\psi_{j-1} + \psi_j]) & 1 < j < J \\
        1 - F_\mathrm{belief}\qty(\frac{1}{2}\qty[\psi_{J-1} + \psi_J])                                                             & j = J.
    \end{cases}
\end{equation}
This step is illustrated in \cref{fig:grid-sketch}.
For higher dimensional projections, this equation can be extended by partitioning the parameter space into a Voronoi diagram, and then integrating the \gls{pdf} $p_\mathrm{belief}$ over each region.
This approach is widely used in analysis of tropical cyclones: first an ensemble of cyclones is simulated, and then probabilistic weights are calculated for each based on attributes such as intensity, landfall location, and radius \citep{johnson_clara:2013,resio_probabilities:2007,toro_jpm-os:2010}.
From a practical perspective, many simulation models round outputs, meaning that there is a possibility of having $\psi_j$ that are equal; in this case it may be helpful to add a small amount of noise to the $\psi_j$ before computing the weights.

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{grid-sketch}
    \caption{
        Schematic of \gls{sow} weighting scheme defined in \cref{eq:weights}.
        This method is illustrated for a hypothetical target distribution (orange line) and $J=4$ samples $\psi_1, \psi_2, \psi_3, \psi_4$ (orange dots).
        As shown in \cref{eq:weights}, the weights $w_j$ (vertical lines) are calculated based on the cumulative distribution function of the target distribution  (black line) at the halfway points $\frac{1}{2}\qty[\psi_i+\psi_{i+1}]$ (vertical dashed lines).
    }\label{fig:grid-sketch}
\end{figure}

The aim of this re-weighting framework is to integrate an ensemble of \glspl{sow} used for exploratory modeling into formal decision analysis.
As in \cref{sec:analysis-condition}, we must condition on a model: where the analysis of \cref{sec:analysis-condition} conditions upon deep uncertainties, the approach outlined in this subsection synthesizes across them.
It is crucially important to understand that stakeholders and experts will not, in general, agree on $p_\mathrm{belief}$ because its uncertainty is epistemic -- that is, there is no ``true'' value of $p_\mathrm{belief}$ that could be estimated if only we had better models and data.
However, reasoning using subjective probability allows\ldots\james{Finish sentence}

@@

We model a one-time decision of whether to elevate a house, and if so by how much (\cref{fig:xlrm}).
We focus on a case study of a \emph{hypothetical} house in Norfolk, VA.
For interpretability, we focus on deep uncertainty in \gls{msl} and treat other model parameters as more shallow uncertainties as shown in \cref{tab:uncertainties}.
We use the notation developed in the previous section to describe the case study.
Specifically,
\begin{enumerate}
    \item The decision vector $\vb{x}$ is a scalar describing how high to elevate a house ($\Delta h$); we consider $\Delta h = \qty{\SI{0}{ft}, \SI{0.25}{ft}, \ldots, \SI{12}{ft}}$ f
    \item The \glspl{sow} describe annual time series of \gls{msl} over the $T=71$ year house lifetime so $\vb{s} \in \mathbb{R}^T$
    \item The system model $f$ quantifies up-front costs and lifetime expected damages, given a decision $x_i$ and \gls{sow} $s_j$, by integrating economic and engineering damage models over a probability distribution for storm surge.
\end{enumerate}
In the remainder of this section we describe data sources and treatment of \gls{slr} (\cref{sec:case-slr}), storm surge (\cref{sec:case-surge}), damages and metrics (\cref{sec:case-metrics}), and finally the subjective priors $p_\mathrm{belief}$ used to apply the re-weighting method described in \cref{sec:analysis-synthesize} to this case study (\cref{sec:case-priors}).

\begin{table}
    \centering
    \caption{
        Summary of parameters, their notation, and how their uncertainty is represented.
        Symbols describing the decision-analytic framework are described in \cref{fig:flowchart}.
    }\label{tab:uncertainties}
    \footnotesize
    \begin{tabular}{p{1.25in} p{0.75in} p{3in}}
        \toprule
        Name                 & Symbol            & Uncertainty                                                                          \\
        \midrule
        \Gls{msl}            & $\overline{y}(t)$ & Deeply uncertain: four physical models $\times$ four \acrshort{rcp} scenarios        \\
        Storm surge          & $y'(t)$           & Probabilistic: Bayesian inference on a stationary \acrshort{gev} model               \\
        Annual maximum flood & $y(t)$            & Deterministic: $y(t)=\overline{y}(t)+y'(t)$                                          \\
        Discount rate        & $1-\gamma$        & Fixed at 2.5\% per year                                                              \\
        Depth-damage         & $D(h-y)$          & Deterministic: based on HAZUS model \citep[see][]{zarekarizi_suboptimal:2020}        \\
        Elevation cost       & $C(\Delta h)$     & Deterministic: a piecewise linear model following \citet{zarekarizi_suboptimal:2020} \\
        Initial height       & $h_0$             & Deterministic: \SI{1}{ft} below the \gls{bfe}, unless otherwise noted                \\
        House floor area     &                   & Deterministic: \SI{1500}{ft^2}                                                       \\
        Structural value     &                   & Deterministic: \usd{200000}                                                          \\
        House lifespan       & $T$               & Deterministic: 71 years                                                              \\
        \bottomrule
    \end{tabular}
    %TODO: check if there are other symbols used
\end{table}

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{xlrm.pdf}
    \caption{
        Conceptual diagram of the considered example.
        A \acrfull{sow} consists of a description of the uncertain factors (red).
        We model a problem with a single lever (yellow), which is how high to elevate a house ($\Delta h$).
        For each \acrshort{sow} and each value of $\Delta h$, the system model (blue) is used to calculate performance metrics (orange).
    }\label{fig:xlrm}
\end{figure}

@@

We analyze simulations of \gls{msl} at Sewells Point, VA from four probabilistic physical models using data published in \citet{ruckert_coastal:2019}.
The four models considered are (i) the BRICK model (version 0.2) with slow (``BRICK Slow'') and (ii) fast (``BRICK Fast'') ice sheet dynamics \citep{wong_brick0.2:2017}, (iii) the \citet{kopp_probabilistic:2014} model (``K14''), and (iv) the \citet{deconto_antarctica:2016} model (``DP16'').
The \citet{kopp_probabilistic:2014} and \citet{deconto_antarctica:2016} models have a ten year time step, which we linearly interpolate onto a one year time step for consistency.
These four models represent physical processes, particularly of ice sheet dynamics, in different ways, leading to diverging sensitivity of \gls{msl} to forcing.
For a discussion of these model outputs we refer the reader to \citet{ruckert_coastal:2019}.

Estimates of nonstationary \gls{msl} also depend on anthropogenic forcing, which is itself deeply uncertain \citep{ho_scenarios:2019,srikrishnan_probabilistic:2022}.
To sample this uncertainty, we use simulations from each physical model under four \gls{rcp} scenarios, yielding sixteen time-varying distributions of \gls{msl}.

@@

The choices of physical model and \gls{rcp} scenario jointly determine future \gls{msl} $p(\overline{y}|t)$.
\Cref{fig:lsl-evolution}(a) shows the time-varying 90\% credible intervals of \gls{msl} for three representative models.
The divergence between the the best-case (blue) and worst-case (red) models is small in the early 21st century and increases rapidly thereafter.
\Cref{fig:lsl-evolution}(b) shows the \glspl{pdf} of mean sea level in 2100 (dashed vertical line in panel (a)) under each of the sixteen models considered.
We return in \cref{sec:results-conditional} to the problem that these multiple models poses to decision makers.

@@

    \item Robustness metrics and robust decision analyses typically make implicit assumptions about the plausibility of future scenarios

@@

    \item Interacting, sequential decisions\james{Commonly }
